Freezing of a Weeks-Chandler-Anderson Fluid using classical Density Functional Theory

### 2020-10-20 papers/thesis-kirstie/thesis.tex
fix quote marks

In developing a model for a fluid, simplifications can be made regarding the composition, type, shape, and interatomic interaction of the molecules. Rather than dealing with any particular molecule or atom, the model only deals with generic "particles", which may also be referred to as "atoms". The shape of the particles, or atoms will be the simplest shape possible - a sphere. The spheres in the simplest models do not exhibit quantum interference effects as the mass of the spheres, and the temperature are taken to be large enough that the de Broglie wavelength is smaller than the mean distance between the spheres. The fluid can then be treated as a classical fluid where the spheres interact with each other according to their electrostatic interatomic potential.

@@

\noindent As seen in Figure \ref{fig:WCA_potential}, the WCA potential decreases from infinity as the distance between the centers of the atoms r increases, and becomes, and remains, essentially zero when r=$2^\frac{1}{6}\sigma=2R\approx{1.122}\sigma$ where R is the radius of the sphere as shown in Figure \ref{fig:TwoSpheres}. As with the Lenard-Jones potential, the repulsive WCA potential increases rapidly and approches infinity as the two spheres begin to overlap. This is a more realistic repulsive potential than that of the Hard Sphere model which goes instantly to infinity when the two spheres make contact (at r=2R), and is zero otherwise. Because the WCA spheres have no hard boundary in their interatomic potential, they are called "soft spheres" and their gradually fading potential makes them slightly "squishy".

@@

For pure substances, a phase transition occurs along a line on a Pressure-Temperature (P-T) phase diagram, such as can be seen in the phase diagram shown in Figure \ref{fig:P-T_Diagram}. This phase diagram is representative of some arbitrary, unknown substance composed of atoms, or molecules which exibit both attraction and repulsion. However, since the atoms in a WCA fliud do not exhibit attraction there will not be a dividing line between the gas and liquid phases, and so the term "liquid" or "gas" is not used, but simply "fluid". Points along the line that separates the solid and the fluid phases indicate a phase transition, and show that the solid and the fluid both have the same temperature and the same pressure at a phase transistion.

@@

The minimization can be illustrated by considering that according to the second law of thermodynamics, a system that is not in equilibrium will exprience spontanteous processes that increase the total entropy of a system and its surroundings until equilibrium is reached, $\Delta{S}_{System + Surroundings} \geq 0$. Here "surroundings" means the entire universe minus the system. This, together with the conservation of energy (the first law of thermodynamics) which says that energy lost (or gained) by the system is gained (or lost) by the surroundings $\Delta{U}_{System}=-\Delta{U}_{Surroundings}$, gives the result that changes in the free energy will be negative for spontaneous processes when the natural variables of the free energy are held fixed. %And so the free energy tends toward a minimum. 
%How will a system at a fixed temperature, volume, and number of particles respond to a small offset from equilibrium?

@@

This cell is rather complicated for the purposes of this discussion, so consider instead a simplified crystal cell with one atom in its center as shown in Figure \ref{fig:simplecell}. 
The Helmholtz free energy is minimized for a fixed volume (and temperature and number of particles).  A fixed cubical volume called the "bold-box" volume, $V_{bold-box}$ will be used for illustration. The number density n for a collection of simplified cells is simply 1 atom per fixed bold-box volume.
\begin{displaymath}{n = \frac{N}{V}=\frac{1~\text{atom}}{V_{bold-box}}}\end{displaymath} It is easier to think of 1 atom per "box" (a reduced number density) than 1 atom per some length$^3$ (the actual number density). Reduced number density is given by n* and is equal to the number of atoms in a "box" of size $\sigma^3$.
\begin{displaymath}{n*=n\sigma^3}\end{displaymath} If $V_{bold-box}$  has a volume of $\sigma^3$, the reduced number density is just the number of atoms in the fixed bold-box volume. 
\begin{displaymath}\text{Reduced Number Density of the simple cell = 1 atom}\end{displaymath}

@@

It is important to note that the number density does not change when introducing a number density profile. There is still a total of one atom per "bold box" volume.

@@

Once again, it is important to note that the number density is the same for all these number density profiles $n(\vec{r})$. There is still a total of one atom per "bold box" volume. This prodcedure is used only to find the crystal free energy as number density profile for a liquid is homogeneous, as shown in Figure \ref{fig:homogen_denisty}.

@@

%------------------n(r)--move to "back" later---------------------
------------------------------------------------------------------

@@

%-------------Minimization-----------------
%The minimization can be shown by considering that according to the second law of thermodynamics, a system that is not in equilibrium will exprience spontanteous processes that increase the total entropy of a system and its surroundings until equilibrium is reached, $\Delta{S}_{System + Surroundings} \geq 0$. Here "surroundings" means the entire universe minus the system. This, together with the conservation of energy (the first law of thermodynamics) which says that whatever energy is lost (or gained) by the system is gained (or lost) by the surroundings $\Delta{U}_{System}=-\Delta{U}_{Surroundings}$, gives the result that any change in the free energy will be negative with each spontaneous process until a minimum is reached if the natural variables of the free energy are held fixed.  How will a system at a fixed temperature, volume, and number of particles respond to a small offset from equilibrium? Rewriting Eq~\ref{usetoshowmin} to show the system terms explicitly
%\begin{equation}d\left(\frac{F}{N}\right)_{System}=d\left(\frac{U}{N}\right)_{System}-Td\left(\frac{S}{N}\right)_{System}+\frac{S}{N}dT_{System}\end{equation}
%and holding the temperature fixed ($T_{System}=T_{Surroundings}$) so that $dT=0$ gives
%\begin{equation}d\left(\frac{F}{N}\right)_{System}=-d\left(\frac{U}{N}\right)_{Surroundings}-Td\left(\frac{S}{N}\right)_{System}\end{equation}
%where conservation of energy was also applied. Fixing the volume and the number of particles, so that $dV=0$ and $d\left(\frac{1}{N}\right)=0$, results in
%\begin{equation}d\left(\frac{F}{N}\right)_{System}=-\frac{T}{N}dS_{System+Surroundings} \end{equation}which, for fixed N, is the same as
%\begin{equation}dF_{System}=-TdS_{System+Surroundings} \end{equation} Since $dS_{System+Surroundings} \geq{0}$, and the temperature and number of particles are fixed, the change in the free energy of the system will always be negative. (reference schroder)

>>>>

In developing a model for a fluid, simplifications can be made regarding the composition, type, shape, and interatomic interaction of the molecules. Rather than dealing with any particular molecule or atom, the model only deals with generic ``particles", which may also be referred to as ``atoms''. The shape of the particles, or atoms will be the simplest shape possible - a sphere. The spheres in the simplest models do not exhibit quantum interference effects as the mass of the spheres, and the temperature are taken to be large enough that the de Broglie wavelength is smaller than the mean distance between the spheres. The fluid can then be treated as a classical fluid where the spheres interact with each other according to their electrostatic interatomic potential.

@@

\noindent As seen in Figure \ref{fig:WCA_potential}, the WCA potential decreases from infinity as the distance between the centers of the atoms r increases, and becomes, and remains, essentially zero when r=$2^\frac{1}{6}\sigma=2R\approx{1.122}\sigma$ where R is the radius of the sphere as shown in Figure \ref{fig:TwoSpheres}. As with the Lenard-Jones potential, the repulsive WCA potential increases rapidly and approches infinity as the two spheres begin to overlap. This is a more realistic repulsive potential than that of the Hard Sphere model which goes instantly to infinity when the two spheres make contact (at r=2R), and is zero otherwise. Because the WCA spheres have no hard boundary in their interatomic potential, they are called ``soft spheres'' and their gradually fading potential makes them slightly ``squishy''.

@@

For pure substances, a phase transition occurs along a line on a Pressure-Temperature (P-T) phase diagram, such as can be seen in the phase diagram shown in Figure \ref{fig:P-T_Diagram}. This phase diagram is representative of some arbitrary, unknown substance composed of atoms, or molecules which exibit both attraction and repulsion. However, since the atoms in a WCA fliud do not exhibit attraction there will not be a dividing line between the gas and liquid phases, and so the term ``liquid" or ``gas" is not used, but simply ``fluid". Points along the line that separates the solid and the fluid phases indicate a phase transition, and show that the solid and the fluid both have the same temperature and the same pressure at a phase transistion.

@@

The minimization can be illustrated by considering that according to the second law of thermodynamics, a system that is not in equilibrium will exprience spontanteous processes that increase the total entropy of a system and its surroundings until equilibrium is reached, $\Delta{S}_{System + Surroundings} \geq 0$. Here ``surroundings" means the entire universe minus the system. This, together with the conservation of energy (the first law of thermodynamics) which says that energy lost (or gained) by the system is gained (or lost) by the surroundings $\Delta{U}_{System}=-\Delta{U}_{Surroundings}$, gives the result that changes in the free energy will be negative for spontaneous processes when the natural variables of the free energy are held fixed. %And so the free energy tends toward a minimum. 
%How will a system at a fixed temperature, volume, and number of particles respond to a small offset from equilibrium?

@@

This cell is rather complicated for the purposes of this discussion, so consider instead a simplified crystal cell with one atom in its center as shown in Figure \ref{fig:simplecell}. 
The Helmholtz free energy is minimized for a fixed volume (and temperature and number of particles).  A fixed cubical volume called the ``bold-box" volume, $V_{bold-box}$ will be used for illustration. The number density n for a collection of simplified cells is simply 1 atom per fixed bold-box volume.
\begin{displaymath}{n = \frac{N}{V}=\frac{1~\text{atom}}{V_{bold-box}}}\end{displaymath} It is easier to think of 1 atom per ``box" (a reduced number density) than 1 atom per some length$^3$ (the actual number density). Reduced number density is given by n* and is equal to the number of atoms in a ``box" of size $\sigma^3$.
\begin{displaymath}{n*=n\sigma^3}\end{displaymath} If $V_{bold-box}$  has a volume of $\sigma^3$, the reduced number density is just the number of atoms in the fixed bold-box volume. 
\begin{displaymath}\text{Reduced Number Density of the simple cell = 1 atom}\end{displaymath}

@@

It is important to note that the number density does not change when introducing a number density profile. There is still a total of one atom per ``bold box" volume.

@@

Once again, it is important to note that the number density is the same for all these number density profiles $n(\vec{r})$. There is still a total of one atom per ``bold box" volume. This prodcedure is used only to find the crystal free energy as number density profile for a liquid is homogeneous, as shown in Figure \ref{fig:homogen_denisty}.

@@

%------------------n(r)--move to ``back" later---------------------
------------------------------------------------------------------

@@

%-------------Minimization-----------------
%The minimization can be shown by considering that according to the second law of thermodynamics, a system that is not in equilibrium will exprience spontanteous processes that increase the total entropy of a system and its surroundings until equilibrium is reached, $\Delta{S}_{System + Surroundings} \geq 0$. Here ``surroundings" means the entire universe minus the system. This, together with the conservation of energy (the first law of thermodynamics) which says that whatever energy is lost (or gained) by the system is gained (or lost) by the surroundings $\Delta{U}_{System}=-\Delta{U}_{Surroundings}$, gives the result that any change in the free energy will be negative with each spontaneous process until a minimum is reached if the natural variables of the free energy are held fixed.  How will a system at a fixed temperature, volume, and number of particles respond to a small offset from equilibrium? Rewriting Eq~\ref{usetoshowmin} to show the system terms explicitly
%\begin{equation}d\left(\frac{F}{N}\right)_{System}=d\left(\frac{U}{N}\right)_{System}-Td\left(\frac{S}{N}\right)_{System}+\frac{S}{N}dT_{System}\end{equation}
%and holding the temperature fixed ($T_{System}=T_{Surroundings}$) so that $dT=0$ gives
%\begin{equation}d\left(\frac{F}{N}\right)_{System}=-d\left(\frac{U}{N}\right)_{Surroundings}-Td\left(\frac{S}{N}\right)_{System}\end{equation}
%where conservation of energy was also applied. Fixing the volume and the number of particles, so that $dV=0$ and $d\left(\frac{1}{N}\right)=0$, results in
%\begin{equation}d\left(\frac{F}{N}\right)_{System}=-\frac{T}{N}dS_{System+Surroundings} \end{equation}which, for fixed N, is the same as
%\begin{equation}dF_{System}=-TdS_{System+Surroundings} \end{equation} Since $dS_{System+Surroundings} \geq{0}$, and the temperature and number of particles are fixed, the change in the free energy of the system will always be negative. (reference schroder)
